Automated Verification for Real-Time Systems using Implicit Clocks and an Extended Antimirov Algorithm

### 2022-04-16 paper/tv.tex
m

%% Automated Temporal Verification for Mixed Sync-Async Reactive Paradigm

@@

\figref{fig:overview_ficher} presents the classical 
Fischer's mutual exclusion algorithm. 
where \codem{d1} and \codem{d2} are two integer constants
 with \codem{d1<d2}; 
\code{x} and \code{ct} are global variables. The protocol is 
modeled as 
process protocol, which is the parallel composition of three 
processes. Each of the three processes attempts to enter 
the critical section when \code{x} is \code{-1}, i.e. no 
other process is 
currently attempting. Once the process is active, it sets \codeme{}{x} 
to its identity \codem{i} within \codem{d1} time units (captured by \codem{deadline(..., d1}). 
Then it idles for \codem{d2} time units (captured by \codem{delay(d2)}) and then 
checks whether \codem{x} is still \codem{i}. If so, 
it enters the critical 
section and leaves later. Otherwise, it restarts from the 
beginning.

We show the specification of \codem{main} using \timedEffects, which
restricts the relation between \codem{d1} and \codem{d2}. When 
\codem{d1<d2} is provided, it ensures a repeated trace where the value of 
\codem{ct} is either \codem{-1} or equals to \codem{x}. The specification
implies that this implementation is indeed mutual exclusive. 
Our prototype system is able to soundly check such time-critical 
algorithms. \autoref{sec:Evaluation} presents the details about the implementation and evaluation.

@@

A transition of the system is of the form $\zeta \xrightarrow[]{l} \zeta^\prime$ 
where $\zeta$ and $\zeta^\prime$ are 
the system configurations before and after the transition respectively. 
%There are four type of the transition labels  :
We adopt the following naming convention \cite{DBLP:conf/icfem/SunLDZ09} for the transition labels \code{l}: 
(1) $t$, denoting a non-negative real integer; 
(2) $\tau$, denoting an invisible event; % which do not make changes to the global valuation;
(3) $\anyevent{A}$ is an observable event. 
%and (4) $\sigma \in \Sigma\cup \{\tau\}$. 
For example, $\zeta \xrightarrow[]{\text{t}} \zeta^\prime$ denotes a transition 
of \code{t} time-units elapsing. 
In the following, we present the firing rules which are associated with 
the timed process constructs. 
{{\small
\begin{flalign*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   Wait                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\code{\frac{
  t {\leq} d 
}{(\env, \textbf{delay}[d]) \xrightarrow[]{\text{t}} (\env, \textbf{delay} [d \text{-}t])}\ [delay_1]} 
\qquad \qquad 
\code{\frac{
}{(\env, \textbf{delay} [0]) \xrightarrow[]{\tau} (\env, ())}\ [delay_2]} 
\end{flalign*}}}
The above captures behaviors of process \textbf{delay}[d]. 
Rule \code{[delay_1]} states that the process may idle for 
any amount of time as long as it is less than or equal to \code{d} time units; 
Rule \code{[delay_2]} states that the process terminates immediately after 
\code{d} becomes \code{0}.
{{\small\begin{flalign*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   timeout                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\code{\frac{
  (\env,e_1) \xrightarrow[]{\anyevent{A}} (\env^\prime,e_1^\prime)
}{(\env, e_1\ \textbf{timeout}  [d] \ e_2) \xrightarrow[]{\anyevent{A}} (\env^\prime,e^\prime)}\ [tot_1]} 
\quad  
\code{\frac{
  (\env,e_1) \xrightarrow[]{\tau} (\env^\prime,e_1^\prime)  
}{(\env, e_1\ \textbf{timeout}  [d] \ e_2) \xrightarrow[]{\text{t}} (\env^\prime, e_1^\prime\ \textbf{timeout}[d]\ e_2 )}\ [tot_2]} 
\end{flalign*}}}

{{\small\begin{flalign*}
\code{\frac{
  (\env,e_1) \xrightarrow[]{\text{t}} (\env,e_1^\prime)  \quad (t {\leq} d) 
}{(\env, e_1\ \textbf{timeout}  [d] \ e_2) \xrightarrow[]{\text{t}} (\env,e_1^\prime\ \textbf{timeout} \  [d\text{-}t] \ e_2)}\ [tot_3]} 
\quad  
\code{\frac{
}{(\env, e_1\ \textbf{timeout}   [0] \ e_2) \xrightarrow[]{\tau} (\env, e_2)}\ [tot_4]} 
\end{flalign*}}}
If an observable event \anyevent{A} can be engaged by \code{e_1}, 
then $e_1\ \textbf{timeout}[d]\ e_2$ becomes \code{e_1^\prime} (rule \code{[tot_1]}). 
An invisible transition continues the waiting on \code{e_1} (rule \code{[tot_2]}). 
If \code{e_1} may idle for less than or equal 
to \code{d} time units, so is the composition (rule \code{[tot_3]}). 
When \code{d} becomes \code{0}, 
\code{e_2} takes over control by a silent transition (rule \code{[tot_4]}).

{{\small\begin{flalign*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%   deadline                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\code{\frac{
  (\env,e) \xrightarrow[]{\anyevent{A}} (\env^\prime,e^\prime)
}{(\env, \textbf{deadline} [d] \ e ) \xrightarrow[]{\anyevent{A}} (\env^\prime, \textbf{deadline}  [d] \ e^\prime)} [ddl_1]} 
\ \ \  
\code{\frac{
  (\env,e) \xrightarrow[]{\text{t}} (\env,e^\prime)  \quad (t {\leq} d)
}{(\env, \textbf{deadline}  [d]\ e ) \xrightarrow[]{\text{t}} (\env, \textbf{deadline}  [d \text{-}t] \ e^\prime )} [ddl_2]} 
\end{flalign*}}}

@@

{{\small\begin{flalign*}
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
  %%   deadline                 %%
  %%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \code{\frac{
    \env \models  b 
  }{(\env, [b] \ e ) \xrightarrow[]{\tau} (\env, e)}\ [guard_1]} 
  \ \ \qquad\qquad 
  \code{\frac{
    \env \not\models  b 
  }{(\env, [b] \ e ) \xrightarrow[]{t} (\env, [b] \ e)}\ [guard_2]} 
\end{flalign*}}}
The guarded process \code{[b]e} behaves as \code{e} when \code{b} is true, 
otherwise it idles until \code{b} becomes true.

{{\small\begin{flalign*}
\code{\frac{
  (\env,e_1) \xrightarrow[]{l} (\env^\prime,e_1^\prime) 
}{(\env, e_1; e_2 ) \xrightarrow[]{l} (\env^\prime, e_1^\prime ; e_2)}\ [seq_1]} 
\qquad
\code{\frac{
}{(\env, (); e_2 ) \xrightarrow[]{\tau} (\env, e_2)}\ [seq_2]} 
\qquad
\code{\frac{
  (\env,e_1) \xrightarrow[]{l} (\env^\prime,e_1^\prime) 
}{(\env, e_1 || e_2 ) \xrightarrow[]{l} (\env^\prime, e_1^\prime || e_2)}\ [par_1]} 
\\ 
  \code{\frac{
    (\env,e_2) \xrightarrow[]{l} (\env^\prime,e_2^\prime) 
  }{(\env, e_1 || e_2 ) \xrightarrow[]{l} (\env^\prime, e_1 || e_2^\prime)}\ [par_2]} 
  \qquad \qquad 
  \code{\frac{
    (\env,e_1) \xrightarrow[]{t} (\env,e_1^\prime) 
    \qquad 
    (\env,e_2) \xrightarrow[]{t} (\env,e_2^\prime) 
  }{(\env, e_1 || e_2 ) \xrightarrow[]{t} (\env, e_1^\prime || e_2^\prime)}\ [par_3]} 
\end{flalign*}}}
Rules \code{[seq_1]} and \code{[seq_2]} state that \code{e_1} takes the control 
when it still can behave; then the control transfers to \code{e_2} when \code{e_1}
terminates. Rules \code{[par_1]} and \code{[par_2]} state that in a parallel composition, 
if any of \code{e_1} or \code{e_2} can proceed, they proceed on their own. Rule 
\code{[par_3]} states that if both of the branches idle, they idle together.

@@

\anyevent{A}\code{(v, \alpha^*)} stands for an observable event with label 
\anyevent{A}, parameterized by \code{v}, and the assignment operations \code{\alpha^*}. 
\code{\tau(\pi)} is an invisible event, which can be parameterized with a pure formula 
\code{\pi}\footnote{The difference between 
\code{\tau(\pi)} and \code{\pi?} is \code{\tau(\pi)} leads to false (\code{\bot}) if 
\code{\pi} is not satisfied, whereas  \code{\pi?} waits until \code{\pi} is satisfied.}. 
To make the specification more expressive, events can also be 
{\anynotevent{A}}, referring to those events which are not 
labeled using \anyevent{A}; and a wildcard \code{\_}, which matches to all the events.

@@

To define the model, 
\code{var} is the set of program variables, 
\code{val} is the set of primitive values, 
\code{\es} is the set of event sequences (or event multi-trees, per se), 
indicating the sequencing constraints on temporal behaviors.
Let \code{d, s, \varphi \models \effect} denote the model relation, i.e., 
the effects \code{\effect} take \code{d} time units to complete; 
the stack \code{s} and linear temporal events \code{\varphi} satisfy the 
temporal effects \code{\effect}, with \code{d}, \code{s}, \code{\varphi} 
from the following concrete domains: \code{d}  {\defeq}\  \code{\mathbb{N}}, 
\code{s}  \defeq\  \code{var {\rightarrow} val } and \code{\varphi}   \defeq\ \code{\es} \ .

@@

@@

\begin{wrapfigure}{R}{0.54\columnwidth}
  \vspace{-3mm}

@@

\vspace{1mm}
$[FV\text{-}\m{Value}]$  obtains the next program state by 
inheriting the current program state.
$[FV\text{-}Guard]$ computes the effects of \code{e} after concatenating
\code{(b{=}True)}? in to the current program states.  
{{\small\begin{gather*}
  \code{\frac{
  }
{  \vdash \{  \Pi, \Theta \}\  v \ 
\{ \Pi, \Theta \}} \  
 {[FV\text{-}Value]}} 
 \qquad
 \code{\frac{
  \vdash \{  \Pi, \Theta \cdot (b{=}True)? \}\  e \ 
  \{ \Pi^\prime, \Theta^\prime \}
  }
{  \vdash \{  \Pi, \Theta \}\  [b]\ e \ 
\{  \Pi^\prime, \Theta^\prime \}} \  
 {[FV\text{-}Guard]}} 
\end{gather*}}}

\vspace{1mm}
$[FV\text{-}Seq]$ computes  \code{ \{  \Pi_1, \Theta_1\} } 
from \code{e_1}, 
then further gets  \code{ \{  \Pi_2, \Theta_2 \} } by 
continuously computing the behaviors of \code{e_2}, to be the final state.
$[FV\text{-}Par]$ computes behaviors for  \code{e_1} and \code{e_2} independently,
then parallel merges the effects from these two branches.  
{{\small\begin{gather*}
  \code{\frac{ 
 \vdash \{ \Pi, \Theta\}  e_1  \{  \Pi_1, \Theta_1\} 
\ \  \ 
\vdash \{ \Pi_1, \Theta_1\}  e_2   \{  \Pi_2, \Theta_2 \} 
}{
  \vdash \{  \Pi, \Theta \}\  
  e_1;e_2 \ \{ \Pi_2, \Theta_2
  \}} 
   {[FV\text{-}Seq]}}  
\quad 
  \code{\frac{
 \vdash \{ \Pi, \Theta \}  e_1   \{  \Pi_1, \Theta_1\} 
\ \  \ 
\vdash \{  \Pi, \Theta \}  e_2   \{  \Pi_2, \Theta_2 \} 
    }{
\vdash \{   \Pi, \Theta \}\  e_1 || e_2 \ \{ 
\Pi_1  \wedge \Pi_2 , \Theta_1 || \Theta_2
\}}  
{[FV\text{-}Par]}} 
\end{gather*}}}

\vspace{1mm}
The \code{[FV\text{-}If\text{-}Else]} computes an 
over-approximation of the program states, by adding different constrains into 
different branches. \code{\Pi \wedge v} enforces 
\code{v} into the pure constraints of every traces in the state, same 
for \code{\Pi \wedge \neg v }.
{{\small\begin{gather*}
  \code{\frac{ 
    \begin{matrix}
\vdash \{ \Pi \wedge v, \Theta    \} \ e_1 \ \{  \Pi_1, \Theta_2 \} 
\qquad 
\vdash \{ \Pi \wedge \neg v , \Theta  \} \ e_2 \ 
\{  \Pi_2, \Theta_2 \}  
\qquad (v\ is\ local)
\end{matrix}}
  {\vdash \s{ \Pi, \Theta}\  \textbf{if}\ v\ \textbf{then}\ e_1\ 
  \textbf{else}\ e_2 \ \s{ \Pi_1, \Theta_2} \cup \s{\Pi_2, \Theta_2} } \  
  {[FV\text{-}If\text{-}Else\text{-}Local]}} 
  \end{gather*}}}
  %\\
%\code{\frac{ 
%  \begin{matrix}
%\vdash \{ \Pi, \Theta    \} \ e_1 \ \{ \Pi_1, \Theta_2 \} 
%\qquad 
%\vdash \{ \Pi , \Theta  \} \ e_2 \ 
%\{  \Pi_2, \Theta_2 \}  
%\qquad (v\ is\ global)
%\end{matrix}}
%{\vdash \s{ \Pi, \Theta}\  \textbf{if}\ v\ \textbf{then}\ e_1\ 
%\textbf{else}\ e_2 \ \s{\Pi, v. \Theta_1 \vee \Theta_2 }} \  
%{[FV\text{-}If\text{-}Else\text{-}Global]}}

\vspace{1mm}
Rule \code{[FV\text{-}Event]} simply constructs an new event and 
concatenates it to the current program states. 
Rule \code{[FV\text{-}Delay]} creates a trace \code{\epsilon\mysharp t}
, where \code{t} is fresh, and concatenates it to the current program states, 
together with the additional constraint \code{t{=}d}. 
  {{\small\begin{gather*}
  \code{\frac{
    \Theta^\prime {=} \Theta\cdot \anyevent{A}(v, \alpha^*) 
    }
  {\vdash \s{ \Pi, \Theta} \  
  \textbf{event} [\anyevent{A}(v, \alpha^*)] 
   \ \{\Pi, \Theta^\prime \}} \   {[FV\text{-}Event]}} 
   \qquad
  \code{\frac{
    \Theta^\prime {=} \Theta\cdot \epsilon\mysharp t \quad (t\ is\ fresh) 
    }
  {\vdash \s{ \Pi, \Theta} \  
  \textbf{delay} [d] 
   \ \{ \Pi \wedge  t {=}d,  \Theta^\prime \}} \   {[FV\text{-}Delay]}} 
\end{gather*}}}

\vspace{1mm}
Rule \code{[FV\text{-}Timeout]} computes the effects from \code{e_1} and \code{e_2}
using the starting state \code{\s{\Pi, \epsilon}}. The final states are the union 
of these effects with their corresponding time variables and arithmetic constrains. 
{{\small\begin{gather*}
  \code{\frac{
    \begin{matrix}
      \vdash \{\Pi, \epsilon \}\ e_1 \ \{  \Pi_1, \Theta_1\} 
      \qquad 
      \vdash \{ \Pi, \epsilon \}\ e_2 \ \{  \Pi_2, \Theta_2 \} 
      \\
      \s{\Theta_f, \Pi_f} {=} \s{\Pi_1  \wedge t_1{\leq}d, (\Theta_1\mysharp t)}
      \cup 
  \s{\Pi_2  \wedge t_2{>}d,  (\Theta_2\mysharp t)}   
  \quad (t_1, t_2\ are\ fresh)     
\end{matrix}
  }
  {\vdash \s{ \Pi, \Theta} \  
  e_1\ \textbf{timeout} [d] \  e_2
   \ \s{\Pi_f, \Theta \cdot \Theta_f} } \   {[FV\text{-}Timeout]}} 
\end{gather*}}}

\vspace{1mm}
Similar to \code{[FV\text{-}Timeout]}, \code{[FV\text{-}Ddl]} computes the effects 
from \code{e} and adds an upper time-bound to the result. 
{{\small\begin{gather*}
  \code{\frac{
    \begin{matrix}
      \vdash \{ \Pi, \epsilon\}\ e \ \{  \Pi_1, \Theta_1\} 
  \quad (t\ is\ fresh)     
\end{matrix}
  }
  {\vdash \s{ \Pi, \Theta} \  
  \textbf{deadline}  [d] \ e 
   \ \s{\Pi_1  \wedge t{\leq}d, \Theta  \cdot (\Theta_1\mysharp t)} }   
    {[FV\text{-}Ddl]}} 
  \end{gather*}}}

  \vspace{1mm}
  Rule \code{[FV\text{-}Meth]} initializes the current effects state 
using the declared precondition, accumulates the effects from the method body, 
to obtain \code{ \s{\Pi, \Theta}}; then  it
checks the  inclusion between \code{ \s{\Pi, \Theta}} and the concatenation of 
the precondition and the postcondition\footnote{Here, 
\code{\effect_{post}} only needs to capture the effects from the current method body, excluding the history effects specified in \code{\effect_{pre}}. }.
  {{\small\begin{gather*}
\code{\frac{  \vdash \s{\effect_{pre}}
 \ e \  \{ \Pi, \Theta \} \qquad   \{ \Pi, \Theta \} \CONTAIN  
 \effect_{pre} \cdot \effect_{post}}{  \vdash  \ mn  \ {x^*}\ \{\textbf{req}\ \effect_{pre}\ 
  \textbf{ens}\  \effect_{post}\} \ \{ e \}}   {[FV\text{-}Meth]}}  
  \end{gather*}}}

  \vspace{1mm}
Rule \code{[FV\text{-}Call]} first checks whether the instantiated 
  precondition of callee, \code{[v^*/x^*]\effect_{pre}}, is satisfied by 
  the current program state. When the check is succeeded, the final states are 
  formed by concatenating the  
  instantiated postcondition to the current states. 
  %, in which we use an auxiliary function \code{rev} to reverse the event sequences of effects.
{{\small\begin{gather*}
\code{\frac{
\begin{matrix}
{\code{  mn  \ {x^*}\ \{\textbf{req}\ \effect_{pre}\ \textbf{ens}\  \effect_{post}\} \ \{ e \} \in \mathcal{P}\ \ }}
\\
\text{\code{ \s{ \Pi, \Theta} \CONTAIN [v^*/x^*] \effect_{pre}   
\qquad   \effect_f =  \s{ \Pi, \Theta} \cdot  [v^*/x^*]\effect_{post}}}
\end{matrix}}
{\vdash \s{ \Pi, \Theta} \ mn({v^*})  \ 
\effect_f} \ {[FV\text{-}Call]}}
\end{gather*}}}

\begin{theorem}[Soundness of the Forward Rules]\label{Soundness_forward_rules}
  Given any system configuration \code{\zeta = (\env, e)}, by applying the operational
  semantics rules, 
  if \code{(\env, e) {\rightarrow^*} (\env^\prime, ())} produces concrete 
  execution time \code{d} and event trace \code{\varphi}, and \code{ \vdash  
  \s{\env, \epsilon} e \s{\Pi, \Theta}}, then, 
  \code{\exists (\pi\wedge \es) \in \s{\Pi, \Theta} } such that 
  \code{d, \env^\prime, \varphi  \models (\pi\wedge \es)}. 
  %.

>>>>

@@

\figref{fig:overview_ficher} presents the classical 
Fischer's mutual exclusion algorithm. 
where \codem{d1} and \codem{d2} are two integer constants
 with \codem{d1<d2}; 
\codem{x} and \codem{ct} are global variables. The protocol is 
modeled as 
process protocol, which is the parallel composition of three 
processes. Each of the three processes attempts to enter 
the critical section when \codem{x} is \codem{-1}, i.e. no 
other process is 
currently attempting. Once the process is active, it sets \codeme{}{x} 
to its identity \codem{i} within \codem{d1} time units (captured by \codem{deadline(..., d1}). 
Then it idles for \codem{d2} time units (captured by \codem{delay(d2)}) and then 
checks whether \codem{x} is still \codem{i}. If so, 
it enters the critical 
section and leaves later. Otherwise, it restarts from the 
beginning.

We show the specification of \codem{main} using \timedEffects, which
restricts the relation between \codem{d1} and \codem{d2}. When 
\codem{d1<d2} is provided, it ensures a global state, i.e., the repeated trace, where the value of 
\codem{ct} is either \codem{-1} or equals to \codem{x}. The specification
implies that this implementation is indeed mutual exclusive. 
Our prototype system is able to soundly check such time-critical 
algorithms. \secref{sec:Evaluation} presents the details about the implementation and evaluation.

@@

A transition of the system is of the form $\zeta \xrightarrow[]{l} \zeta^\prime$ 
where $\zeta$ and $\zeta^\prime$ are 
the system configurations before and after the transition respectively. 
%There are four type of the transition labels  :
We adopt the following naming convention \cite{DBLP:conf/icfem/SunLDZ09} for the transition labels \code{l}: 
(1) $t$, denoting a non-negative real integer; 
(2) $\tau$, denoting an invisible event; % which do not make changes to the global valuation;
(3) $\anyevent{A}$ is an observable event. 
%and (4) $\sigma \in \Sigma\cup \{\tau\}$. 
For example, $\zeta \xrightarrow[]{\text{t}} \zeta^\prime$ denotes a transition 
of \code{t} time-units elapsing. 
In the following, we present the firing rules which are associated with 
the timed process constructs.

\input{rules/op_delay.tex}

The above captures behaviors of process \textbf{delay}[d]. 
Rule \code{[delay_1]} states that the process may idle for 
any amount of time as long as it is less than or equal to \code{d} time units; 
Rule \code{[delay_2]} states that the process terminates immediately after 
\code{d} becomes \code{0}.

\input{rules/op_to.tex}

If an observable event \anyevent{A} can be engaged by \code{e_1}, 
then $e_1\ \textbf{timeout}[d]\ e_2$ becomes \code{e_1^\prime} (rule \code{[tot_1]}). 
An invisible transition continues the waiting on \code{e_1} (rule \code{[tot_2]}). 
If \code{e_1} may idle for less than or equal 
to \code{d} time units, so is the composition (rule \code{[tot_3]}). 
When \code{d} becomes \code{0}, 
\code{e_2} takes over control by a silent transition (rule \code{[tot_4]}).

\input{rules/op_ddl.tex}

@@

\input{rules/op_gu.tex}

The guarded process \code{[b]e} behaves as \code{e} when \code{b} is true, 
otherwise it idles until \code{b} becomes true.

{{\small\begin{flalign*}
\code{\frac{
  (\env,e_1) \xrightarrow[]{l} (\env^\prime,e_1^\prime) 
}{(\env, e_1; e_2 ) \xrightarrow[]{l} (\env^\prime, e_1^\prime ; e_2)}\ [seq_1]} 
\qquad
\code{\frac{
}{(\env, (); e_2 ) \xrightarrow[]{\tau} (\env, e_2)}\ [seq_2]} 
\qquad
\code{\frac{
  (\env,e_1) \xrightarrow[]{l} (\env^\prime,e_1^\prime) 
}{(\env, e_1 || e_2 ) \xrightarrow[]{l} (\env^\prime, e_1^\prime || e_2)}\ [par_1]} 
\\ 
  \code{\frac{
    (\env,e_2) \xrightarrow[]{l} (\env^\prime,e_2^\prime) 
  }{(\env, e_1 || e_2 ) \xrightarrow[]{l} (\env^\prime, e_1 || e_2^\prime)}\ [par_2]} 
  \qquad \qquad 
  \code{\frac{
    (\env,e_1) \xrightarrow[]{t} (\env,e_1^\prime) 
    \qquad 
    (\env,e_2) \xrightarrow[]{t} (\env,e_2^\prime) 
  }{(\env, e_1 || e_2 ) \xrightarrow[]{t} (\env, e_1^\prime || e_2^\prime)}\ [par_3]} 
\end{flalign*}}}

Rules \code{[seq_1]} and \code{[seq_2]} state that \code{e_1} takes the control 
when it still can behave; then the control transfers to \code{e_2} when \code{e_1}
terminates. Rules \code{[par_1]} and \code{[par_2]} state that in a parallel composition, 
if any of \code{e_1} or \code{e_2} can proceed, they proceed on their own. Rule 
\code{[par_3]} states that if both of the branches idle, they idle together.

@@

\anyevent{A}\code{(v, \alpha^*)} stands for an observable event with label 
\anyevent{A}, parameterized by \code{v}, and the assignment operations \code{\alpha^*}; and 
\code{\tau(\pi)} is an invisible event, which can be parameterized with a pure formula 
\code{\pi}\footnote{The difference between 
\code{\tau(\pi)} and \code{\pi?} is \code{\tau(\pi)} leads to false (\code{\bot}) if 
\code{\pi} is not satisfied, whereas  \code{\pi?} waits until \code{\pi} is satisfied.}. 
To make the specification more expressive, events can also be 
{\anynotevent{A}}, referring to those events which are not 
labeled using \anyevent{A}; and a wildcard \code{\_}, which matches to all the events.

@@

To define the model, 
\code{var} is the set of program variables, 
\code{val} is the set of primitive values, 
\code{\es} is the set of event sequences (or event multi-trees, per se), 
indicating the sequencing constraints on temporal behaviors.

@@

Let \code{d, s, \varphi \models \effect} denote the model relation, i.e., 
the effects \code{\effect} take \code{d} time units to complete; 
the stack \code{s} and linear temporal events \code{\varphi} satisfy the 
temporal effects \code{\effect}, with \code{d}, \code{s}, \code{\varphi} 
from the following concrete domains: \code{d}  {\defeq}\  \code{\mathbb{N}}, 
\code{s}  \defeq\  \code{var {\rightarrow} val } and \code{\varphi}   \defeq\ \code{\es} \ .

@@

\begin{wrapfigure}{R}{0.54\columnwidth}
 % \vspace{-1mm}

@@

\vspace{1mm}
$[FV\text{-}\m{Value}]$  obtains the next program state by 
inheriting the current program state.
$[FV\text{-}Guard]$ computes the effects of \code{e} after concatenating
\code{(b{=}True)}? in to the current program states.

{{\small\begin{gather*}
  \code{\frac{
  }
{  \vdash \{  \Pi, \Theta \}\  v \ 
\{ \Pi, \Theta \}} \  
 {[FV\text{-}Value]}} 
 \qquad
 \code{\frac{
  \vdash \{  \Pi, \Theta \cdot (b{=}True)? \}\  e \ 
  \{ \Pi^\prime, \Theta^\prime \}
  }
{  \vdash \{  \Pi, \Theta \}\  [b]\ e \ 
\{  \Pi^\prime, \Theta^\prime \}} \  
 {[FV\text{-}Guard]}} 
\end{gather*}}}

\vspace{1mm}
$[FV\text{-}Seq]$ computes  \code{ \{  \Pi_1, \Theta_1\} } 
from \code{e_1}, 
then further gets  \code{ \{  \Pi_2, \Theta_2 \} } by 
continuously computing the behaviors of \code{e_2}, to be the final state.
$[FV\text{-}Par]$ computes behaviors for  \code{e_1} and \code{e_2} independently,
then parallel merges the effects from these two branches.

{{\small\begin{gather*}
  \code{\frac{ 
 \vdash \{ \Pi, \Theta\}  e_1  \{  \Pi_1, \Theta_1\} 
\ \  \ 
\vdash \{ \Pi_1, \Theta_1\}  e_2   \{  \Pi_2, \Theta_2 \} 
}{
  \vdash \{  \Pi, \Theta \}\  
  e_1;e_2 \ \{ \Pi_2, \Theta_2
  \}} 
   {[FV\text{-}Seq]}}  
\quad 
  \code{\frac{
 \vdash \{ \Pi, \Theta \}  e_1   \{  \Pi_1, \Theta_1\} 
\ \  \ 
\vdash \{  \Pi, \Theta \}  e_2   \{  \Pi_2, \Theta_2 \} 
    }{
\vdash \{   \Pi, \Theta \}\  e_1 || e_2 \ \{ 
\Pi_1  \wedge \Pi_2 , \Theta_1 || \Theta_2
\}}  
{[FV\text{-}Par]}} 
\end{gather*}}}

The \code{[FV\text{-}If\text{-}Else]} computes an 
over-approximation of the program states, by adding different constrains into 
different branches. \code{\Pi \wedge v} enforces 
\code{v} into the pure constraints of every traces in the state, same 
for \code{\Pi \wedge \neg v }.

{{\small\begin{gather*}
  \code{\frac{ 
    \begin{matrix}
\vdash \{ \Pi \wedge v, \Theta    \} \ e_1 \ \{  \Pi_1, \Theta_2 \} 
\qquad 
\vdash \{ \Pi \wedge \neg v , \Theta  \} \ e_2 \ 
\{  \Pi_2, \Theta_2 \}  
\qquad (v\ is\ local)
\end{matrix}}
  {\vdash \s{ \Pi, \Theta}\  \textbf{if}\ v\ \textbf{then}\ e_1\ 
  \textbf{else}\ e_2 \ \s{ \Pi_1, \Theta_2} \cup \s{\Pi_2, \Theta_2} } \  
  {[FV\text{-}If\text{-}Else\text{-}Local]}} 
  \end{gather*}}}
  %\\
%\code{\frac{ 
%  \begin{matrix}
%\vdash \{ \Pi, \Theta    \} \ e_1 \ \{ \Pi_1, \Theta_2 \} 
%\qquad 
%\vdash \{ \Pi , \Theta  \} \ e_2 \ 
%\{  \Pi_2, \Theta_2 \}  
%\qquad (v\ is\ global)
%\end{matrix}}
%{\vdash \s{ \Pi, \Theta}\  \textbf{if}\ v\ \textbf{then}\ e_1\ 
%\textbf{else}\ e_2 \ \s{\Pi, v. \Theta_1 \vee \Theta_2 }} \  
%{[FV\text{-}If\text{-}Else\text{-}Global]}}

Rule \code{[FV\text{-}Event]} simply constructs an new event and 
concatenates it to the current program states. 
Rule \code{[FV\text{-}Delay]} creates a trace \code{\epsilon\mysharp t}
, where \code{t} is fresh, and concatenates it to the current program states, 
together with the additional constraint \code{t{=}d}.

  {{\small\begin{gather*}
  \code{\frac{
    \Theta^\prime {=} \Theta\cdot \anyevent{A}(v, \alpha^*) 
    }
  {\vdash \s{ \Pi, \Theta} \  
  \textbf{event} [\anyevent{A}(v, \alpha^*)] 
   \ \{\Pi, \Theta^\prime \}} \   {[FV\text{-}Event]}} 
   \qquad
  \code{\frac{
    \Theta^\prime {=} \Theta\cdot \epsilon\mysharp t \quad (t\ is\ fresh) 
    }
  {\vdash \s{ \Pi, \Theta} \  
  \textbf{delay} [d] 
   \ \{ \Pi \wedge  t {=}d,  \Theta^\prime \}} \   {[FV\text{-}Delay]}} 
\end{gather*}}}

Rule \code{[FV\text{-}Timeout]} computes the effects from \code{e_1} and \code{e_2}
using the starting state \code{\s{\Pi, \epsilon}}. The final states are the union 
of these effects with their corresponding time variables and arithmetic constrains. 
Note that, \code{hd(\Theta_1)} and \code{tl(\Theta_1)} return the head and the tail of 
\code{\Theta_1} respectively.

{{\small\begin{gather*}
  \code{\frac{
    \begin{matrix}
      \vdash \{\Pi, \epsilon \}\ e_1 \ \{  \Pi_1, \Theta_1\} 
      \qquad 
      \vdash \{ \Pi, \epsilon \}\ e_2 \ \{  \Pi_2, \Theta_2 \} 
      \\
      \s{\Theta_f, \Pi_f} {=} \s{\Pi_1  \wedge t_1{\leq}d, (hd(\Theta_1)\mysharp t \cdot tl(\Theta_1))}
      \cup 
  \s{\Pi_2  \wedge t_2{>}d,  (\Theta_2\mysharp t)}   
  \quad (t_1, t_2\ are\ fresh)     
\end{matrix}
  }
  {\vdash \s{ \Pi, \Theta} \  
  e_1\ \textbf{timeout} [d] \  e_2
   \ \s{\Pi_f, \Theta \cdot \Theta_f} } \   {[FV\text{-}Timeout]}} 
\end{gather*}}}

Similar to \code{[FV\text{-}Timeout]}, \code{[FV\text{-}Ddl]} computes the effects 
from \code{e} and adds an upper time-bound to the result.

{{\small\begin{gather*}
  \code{\frac{
    \begin{matrix}
      \vdash \{ \Pi, \epsilon\}\ e \ \{  \Pi_1, \Theta_1\} 
  \quad (t\ is\ fresh)     
\end{matrix}
  }
  {\vdash \s{ \Pi, \Theta} \  
  \textbf{deadline}  [d] \ e 
   \ \s{\Pi_1  \wedge t{\leq}d, \Theta  \cdot (\Theta_1\mysharp t)} }   
    {[FV\text{-}Ddl]}} 
  \end{gather*}}}

  Rule \code{[FV\text{-}Meth]} initializes the current effects state 
using the declared precondition, accumulates the effects from the method body, 
to obtain \code{ \s{\Pi, \Theta}}; then  it
checks the  inclusion between \code{ \s{\Pi, \Theta}} and the concatenation of 
the precondition and the postcondition\footnote{Here, 
\code{\effect_{post}} only needs to capture the effects from the current method body, excluding the history effects specified in \code{\effect_{pre}}. }.

{{\small\begin{gather*}
\code{\frac{  \vdash \s{\effect_{pre}}
 \ e \  \{ \Pi, \Theta \} \qquad   \{ \Pi, \Theta \} \CONTAIN  
 \effect_{pre} \cdot \effect_{post}}{  \vdash  \ mn  \ {x^*}\ \{\textbf{req}\ \effect_{pre}\ 
  \textbf{ens}\  \effect_{post}\} \ \{ e \}}   {[FV\text{-}Meth]}}  
  \end{gather*}}}

Rule \code{[FV\text{-}Call]} first checks whether the instantiated 
  precondition of callee, \code{[v^*/x^*]\effect_{pre}}, is satisfied by 
  the current program state. When the check is succeeded, the final states are 
  formed by concatenating the  
  instantiated postcondition to the current states.

  %, in which we use an auxiliary function \code{rev} to reverse the event sequences of effects.
{{\small\begin{gather*}
\code{\frac{
\begin{matrix}
{\code{  mn  \ {x^*}\ \{\textbf{req}\ \effect_{pre}\ \textbf{ens}\  \effect_{post}\} \ \{ e \} \in \mathcal{P}\ \ }}
\\
\text{\code{ \s{ \Pi, \Theta} \CONTAIN [v^*/x^*] \effect_{pre}   
\qquad   \effect_f =  \s{ \Pi, \Theta} \cdot  [v^*/x^*]\effect_{post}}}
\end{matrix}}
{\vdash \s{ \Pi, \Theta} \ mn({v^*})  \ 
\effect_f} \ {[FV\text{-}Call]}}
\end{gather*}}}

\begin{theorem}[Soundness of the Forward Rules]\label{Soundness_forward_rules}
  Given any system configuration \code{\zeta = (\env, e)}, by applying the operational
  semantics rules, 
  if \code{(\env, e) {\rightarrow^*} (\env^\prime, ())} produces concrete 
  execution time \code{d} and event sequence \code{\varphi}, and the forward verifier reasons \code{ \vdash  
  \s{\env, \epsilon} \ e\  \s{\Pi, \Theta}}, then, \\
  \code{\qquad \qquad \qquad \qquad  \qquad 
   \exists (\pi\wedge \es) \in \s{\Pi, \Theta} } such that 
  \code{d, \env^\prime, \varphi  \models (\pi\wedge \es)}. 
  %.
